Switch equivalence and weak prime domains for left-linear DPO rewriting systems

### 2025-03-21 full_version/leftlinear.tex
minori

\begin{lemma}[Uniqueness of pre-switches]
	\label{thm:switch_uni}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are pre-switches of $\dder{D}_0$ and
	$\dder{D}_1$. Then there is a unique abstraction equivalence $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ between $\der{E}$ and $\der{F}$, whose first component is $\id{G_0}$. Moreover, 
	\[\phi_{D_{\der{E}, 0}}\circ j_1=a_1 \qquad \phi_{D_{\der{E}, 1}}\circ j_0=a_0\]
\end{lemma}

@@

	Now, by \Cref{prop:unique} we get isomorphisms $\phi_0\colon D_{\der{E},0}\to D_{\der{F},0}$ and $\psi_1 \colon G_{\der{E},1}\to G_{\der{F},1}$ such that the diagram below commutes.
	\[\xymatrix{G_{0} & \ar@{>->}[l]_{f_{\der{F},0}} D_{\der{F},0} \ar[r]^{g_{\der{F},0}}& G_{\der{F},1} \\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{h_{\der{E},0}}  \ar[u]^{h_{\der{F},0}}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} \ar@/_.5cm/[uu]_{\psi_1}}\]

@@

	Moreover, we have already shown that $\psi_1\circ m_{\der{E},1}=m_{\der{F},1}$. Thus, by \Cref{prop:unique} there are the dotted isomorphisms $\phi_1\colon D_{\der{E},1}\to D_{\der{F},1}$ and $\psi_2\colon G_2\to G_2$ in the diagram below.
	\[\xymatrix{G_{\der{F},1} & \ar@{>->}[l]_{f_{\der{F},1}} D_{\der{F},1} \ar[r]^{g_{\der{F},1}}& G_2 \\L_0 \ar[d]^{m_{\der{E},1}} \ar[u]_{m_{\der{F},1}}& K_{0} \ar[d]^{k_{\der{E},1}} \ar[u]_{k_{\der{F},1}}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_{0}  \ar[d]_{g_1\circ i_0}  \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\psi_1} G_{\der{E},1} & \ar@{>->}[l]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{1}}|\hole \ar[r]_{g_{\der{E},1}}& G_2 \ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]

@@

\begin{example}\label{ex:abs2}
Let $\der{D}$ be the derivation of \Cref{ex:abs} and consider the pre-switch $\der{F}$. Then we have the following diagram

	\[\xymatrix@C=22pt@R=22pt{&&1 \ar@{>->}@/_.9cm/[ddrr]_(.2){\delta^3_0}|(.72)\hole
	\ar@{>->}[dr]^{\delta^3_0}&& 1\ar@{>->}@/^.9cm/[ddll]^(.2){\delta^3_0}
	\ar@{>->}[dl]_{\delta^3_0}\\
	&1\ar@{>->}[dr]^{\delta^3_0}\ar@{>->}[dl]_{\id{1}}
	\ar@{>->}@/_.7cm/[ddrr]|(.38)\hole^(.7){\delta^3_0}\ar@{>->}[ur]^{\id{1}}&& 3 &&
	1
	\ar@{>->}@/^.7cm/[ddll]|(.38)\hole_(.7){\delta^3_0}\ar@{>->}[dl]_{\delta^3_0}\ar@{>->}[lu]_{\delta^3_0}
	\ar@{>->}[dr]^{\delta^2_0}\\
	1 \ar@/_.7cm/[ddrr]_(.2){\delta^3_0}|(.325)\hole|(.81)\hole
	\ar@{>->}[dr]^(.35){\delta^3_0}|(.6)\hole && 3
	\ar@{>->}[dl]|(.4)\hole^(.65){\id{3}}\ar@{>->}[ur]|(.5)\hole^(.7){\id{3} }&&3
	\ar@{>->}[dr]|(.4)\hole_(.65){i_3}
	\ar@{>->}[ul]|(.475)\hole_(.7){\id{3}}&&2\ar@{>->}@/^.7cm/[ddll]^(.2){i_2}|(.325)\hole|(.81)\hole\ar@{>->}[dl]_(.35){i_2}|(.6)\hole\\
	&3 &&3	\ar@{>->}[dl]^{\id{3}}\ar@{>->}[ur]^{\id{3}}\ar@{>->}[ul]_{\id{3}}&&4\\
	1	\ar@{>->}@/^.7cm/[uurr]^(.2){\delta^3_0} \ar@{>->}[ur]_(.35){\delta^3_0}|(.61)\hole && 3	\ar@{>->}[ul]|(.4)\hole_(.65){\id{3}}\ar@{>->}[dr]|(.5)\hole_(.65){i_3}	&&4\ar@{>->}[ur]|(.4)\hole^(.65){t} \ar@{>->}[dl]|(.5)\hole^(.65){\id{4}}	&& 1\ar[ul]^(.35){\delta^4_0}|(.61)\hole\ar@/_.7cm/[uull]_(.2){\delta^3_0}\\ 
	&1	\ar@{>->}[ur]_{\delta^3_0} \ar@{>->}[dr]_{i_2}	\ar@{>->}[ul]^{\id{1}}\ar@/^.7cm/[uurr]_(.65){\delta^3_0}&&4&& 1\ar@{>->}[ur]_{\id{1}} \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}[dl]^{\id{1}} \ar@{>->}@/_.7cm/[uull]^(.65){\delta^3_0}\\&& 2	\ar@{>->}[ur]_{i_2}\ar@{>->}@/^.9cm/[uurr]^(.2){i_2}&& 1 \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}@/_.9cm/[uull]_(.2){\delta^3_0} |(.72)\hole }\]

@@

\begin{definition}(Switch) 	Let $\X$ be an $\mathcal{M}$-adhesive category and $(\X, \R)$ a left-linear rewriting system. 
	Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram below, where the cone with vertex $P$ is a pullback and $u_0\colon K_0\to P$ and $u_1\colon K_1\to P$ are the arrows built in \Cref{prop:tec}.
	\[\xymatrix@R=20pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0 \ar @/^.35cm/[ddrr]^(.3){u_0}
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}|(.68)\hole
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}|(.68)\hole 
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar @/_.35cm/[ddll]_(.3){u_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}|(.7)\hole&& G_1 && \ar@{>->}[ll]^{f_1}|(.7)\hole D_1
		\ar[rr]_{g_1}&& G_2 \\ & & & & P \ar[urr]_{p_1} \ar@{>->}[ull]^{p_0}}
	\]

	Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$, depicted below, be a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=20pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	We say that $\der{E}$ is a \emph{switch} for $\der{D}$ along $(i_0, i_1)$ if there is an arrow $z_1\colon P\to D'_1$ fitting in the diagram below, where $z_0\colon P\to D'_0$ is the arrow constructed in \Cref{lem:cose}.
	\[\xymatrix{D'_0  \ar[d]_{g'_0}& P \ar[r]^{p_1} \ar[l]_{z_0} \ar@{.>}[d]^{z_1} & D_1 \ar[d]^{g_1}\\ G'_1 & D'_1 \ar[r]_{g'_1} \ar@{>->}[l]_{f'_1} & G_2}\]

	If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we say that they are \emph{switchable}. If $(\der{D}, \alpha, \omega)$ is a decorated derivation of length $2$, and $(i_0, i_1)$ an independence pair between $\dder{D}_0$ and $\dder{D}_1$. A \emph{decorated switch} of $(\der{D}, \alpha, \omega)$ along $(i_0, i_1)$ is $(\der{E}, \alpha, \omega)$, where $\der{E}$ is a switch for $\der{D}$ along $(i_0, i_1)$.
\end{definition}

\begin{remark} 
vari quadrati
\end{remark}

\begin{lemma}
	equivalenze

	\begin{enumerate}
		\item 
		\item 
	\end{enumerate}
\end{lemma}
\begin{proof}
	contenuto...
\end{proof}

We end this section showing that, if a switch exists, then any pre-switch can be turned into one.

>>>>

\begin{lemma}[Uniqueness of pre-switches]
	\label{thm:switch_uni}
	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are pre-switches of $\dder{D}_0$ and
	$\dder{D}_1$ along the same independence pair $(i_0, i_1)$. Then there is a unique abstraction equivalence $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ between $\der{E}$ and $\der{F}$, whose first component is $\id{G_0}$. Moreover, 
	\[\phi_{D_{\der{E}, 0}}\circ j_1=a_1 \qquad \phi_{D_{\der{E}, 1}}\circ j_0=a_0\]
\end{lemma}

@@

	Now, by \Cref{prop:unique} we know that there are two unique isomorphisms $\phi_0\colon D_{\der{E},0}\to D_{\der{F},0}$ and $\psi_1 \colon G_{\der{E},1}\to G_{\der{F},1}$ such that the diagram below commutes.
	\[\xymatrix{G_{0} & \ar@{>->}[l]_{f_{\der{F},0}} D_{\der{F},0} \ar[r]^{g_{\der{F},0}}& G_{\der{F},1} \\L_1 \ar[d]^{f_0\circ i_1} \ar[u]_{f_0\circ i_1}& K_{1} \ar[d]^{k_{\der{E},0}} \ar[u]_{k_{\der{F},0}}\ar[l]_{l_1} \ar[r]^{r_1} & R_{1}  \ar[d]_{h_{\der{E},0}}  \ar[u]^{h_{\der{F},0}}\\ \ar@/^.5cm/[uu]^{\id{G_0}} G_{0} & \ar@{>->}[l]^{f_{\der{E},0}} D_{\der{E},0} \ar@/^.5cm/[uu]^(.33){\phi_{0}}|\hole \ar[r]_{g_{\der{E},0}}& G_{\der{E},1} \ar@/_.5cm/[uu]_{\psi_1}}\]

@@

	Moreover, we have already shown that $\psi_1\circ m_{\der{E},1}=m_{\der{F},1}$. Thus, again by \Cref{prop:unique} there are the dotted isomorphisms $\phi_1\colon D_{\der{E},1}\to D_{\der{F},1}$ and $\psi_2\colon G_2\to G_2$ in the diagram below and such arrows are unique.
	\[\xymatrix{G_{\der{F},1} & \ar@{>->}[l]_{f_{\der{F},1}} D_{\der{F},1} \ar[r]^{g_{\der{F},1}}& G_2 \\L_0 \ar[d]^{m_{\der{E},1}} \ar[u]_{m_{\der{F},1}}& K_{0} \ar[d]^{k_{\der{E},1}} \ar[u]_{k_{\der{F},1}}\ar@{>->}[l]_{l_1} \ar[r]^{r_1} & R_{0}  \ar[d]_{g_1\circ i_0}  \ar[u]^{g_1\circ i_0}\\ \ar@/^.5cm/[uu]^{\psi_1} G_{\der{E},1} & \ar@{>->}[l]^{f_{\der{E},1}} D_{\der{E},1} \ar@{.>}@/^.5cm/[uu]^(.33){\phi_{1}}|\hole \ar[r]_{g_{\der{E},1}}& G_2 \ar@{.>}@/_.5cm/[uu]_{\psi_2}}\]

@@

\begin{example}\label{ex:abs2}
Let $\der{D}$ be the derivation of \Cref{ex:abs} and consider the pre-switch $\der{F}$. Then we can build the diagram belo
	\[\xymatrix@C=20pt@R=20pt{&&1 \ar@{>->}@/_.9cm/[ddrr]_(.2){\delta^3_0}|(.72)\hole
	\ar@{>->}[dr]^{\delta^3_0}&& 1\ar@{>->}@/^.9cm/[ddll]^(.2){\delta^3_0}
	\ar@{>->}[dl]_{\delta^3_0}\\
	&1\ar@{>->}[dr]^{\delta^3_0}\ar@{>->}[dl]_{\id{1}}
	\ar@{>->}@/_.7cm/[ddrr]|(.38)\hole^(.7){\delta^3_0}\ar@{>->}[ur]^{\id{1}}&& 3 &&
	1
	\ar@{>->}@/^.7cm/[ddll]|(.38)\hole_(.7){\delta^3_0}\ar@{>->}[dl]_{\delta^3_0}\ar@{>->}[lu]_{\delta^3_0}
	\ar@{>->}[dr]^{\delta^2_0}\\
	1 \ar@/_.7cm/[ddrr]_(.2){\delta^3_0}|(.325)\hole|(.81)\hole
	\ar@{>->}[dr]^(.35){\delta^3_0}|(.6)\hole && 3
	\ar@{>->}[dl]|(.4)\hole^(.65){\id{3}}\ar@{>->}[ur]|(.5)\hole^(.7){\id{3} }&&3
	\ar@{>->}[dr]|(.4)\hole_(.65){i_3}
	\ar@{>->}[ul]|(.475)\hole_(.7){\id{3}}&&2\ar@{>->}@/^.7cm/[ddll]^(.2){i_2}|(.325)\hole|(.81)\hole\ar@{>->}[dl]_(.35){i_2}|(.6)\hole\\
	&3 &&3	\ar@{>->}[dl]^{\id{3}}\ar@{>->}[ur]^{\id{3}}\ar@{>->}[ul]_{\id{3}}&&4\\
	1	\ar@{>->}@/^.7cm/[uurr]^(.2){\delta^3_0} \ar@{>->}[ur]_(.35){\delta^3_0}|(.61)\hole && 3	\ar@{>->}[ul]|(.4)\hole_(.65){\id{3}}\ar@{>->}[dr]|(.5)\hole_(.65){i_3}	&&4\ar@{>->}[ur]|(.4)\hole^(.65){t} \ar@{>->}[dl]|(.5)\hole^(.65){\id{4}}	&& 1\ar[ul]^(.35){\delta^4_0}|(.61)\hole\ar@/_.7cm/[uull]_(.2){\delta^3_0}\\ 
	&1	\ar@{>->}[ur]_{\delta^3_0} \ar@{>->}[dr]_{i_2}	\ar@{>->}[ul]^{\id{1}}\ar@/^.7cm/[uurr]_(.65){\delta^3_0}&&4&& 1\ar@{>->}[ur]_{\id{1}} \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}[dl]^{\id{1}} \ar@{>->}@/_.7cm/[uull]^(.65){\delta^3_0}\\&& 2	\ar@{>->}[ur]_{i_2}\ar@{>->}@/^.9cm/[uurr]^(.2){i_2}&& 1 \ar@{>->}[ul]^{\delta^4_0} \ar@{>->}@/_.9cm/[uull]_(.2){\delta^3_0} |(.72)\hole }\]

@@

\begin{definition}[Switch] 	Let $(\X, \R)$ be a left-linear rewriting system, with $\X$ an $\mathcal{M}$-adhesive category.
	Let also $\der{D}=\{\dder{D}_i\}_{i=0}^1$ be a derivation made
	of two sequentially independent derivations $\dder{D}_0$,
	$\dder{D}_1$ with independence pair $(i_0, i_1)$, as in the diagram below, where the cone with vertex $P$ is a pullback and $u_0\colon K_0\to P$ and $u_1\colon K_1\to P$ are the arrows built in \Cref{prop:tec}.
	\[\xymatrix@R=15pt@C=20pt{L_0 \ar[d]_{m_0}&& K_0 \ar @/^.35cm/[ddrr]_(.2){u_0}
		\ar[d]_{k_0}\ar@{>->}[ll]_{l_0} \ar[r]^{r_0} & R_0
		\ar@{>}@/^.25cm/[drrr]|(.32)\hole_(.4){i_0}|(.68)\hole
		\ar[dr]|(.33)\hole_{h_0} && L_1 \ar@{>}@/_.25cm/[dlll]^(.4){i_1}|(.68)\hole 
		\ar[dl]|(.33)\hole^{m_1}& K_1 \ar[d]^{k_1}\ar@{>->}[l]_{l_1} \ar @/_.35cm/[ddll]^(.2){u_1}
		\ar[rr]^{r_1} && R_1 \ar[d]^{h_1}\\G_0 && \ar@{>->}[ll]^{f_0}
		D_0 \ar[rr]_{g_0}|(.7)\hole&& G_1 && \ar@{>->}[ll]^{f_1}|(.7)\hole D_1
		\ar[rr]_{g_1}&& G_2 \\ & & & & P \ar[urr]_{p_1} \ar@{>->}[ull]^{p_0}}
	\]

	Let $\der{E}=\{\dder{E}_i\}_{i=0}^1$, depicted below, be a pre-switch for $\der{D}$ along $(i_0,i_1)$.
	\[
	\xymatrix@R=18pt@C=20pt{
		{L_1} \ar[d]_{m_0'}
		&&  {K_1} \ar@{>->}[ll]_{l_1} \ar[r]^{r_1} \ar[d]_{k_0'}
		&  {R_1} \ar[dr]|(.33)\hole_{h_0'}  \ar@/^.25cm/@{>}_(.4){i_0'}|(.32)\hole[drrr]
		& & 
		{L_0}\ar[dl]|(.33)\hole^{m_1'} \ar@/_.25cm/@{>}^(.4){i_1'}[dlll] 
		&  {K_0} \ar@{>->}[l]_{l_0} \ar[rr]^{r_0} \ar[d]^{k_1'}
		& & {R_0} \ar[d]^{h_1'} \\		
		{G_0}
		& & {D_0'} \ar@{>->}[ll]^{f_0'} \ar[rr]_{g_0'}
		& &  {G'_1} 
		& &  {D_1'} \ar@{>->}[ll]^{f_1'} \ar[rr]_{g_1'}
		& & {G_2}  }
	\]

	We say that $\der{E}$ is a \emph{switch} for $\der{D}$ along $(i_0, i_1)$ if there is an arrow $z_1\colon P\to D'_1$ fitting in the diagram below, where $z_0\colon P\to D'_0$ is the arrow constructed in \Cref{lem:cose}.
	\[\xymatrix@R=18pt{D'_0  \ar[d]_{g'_0}& P \ar[r]^{p_1} \ar[l]_{z_0} \ar@{.>}[d]^{z_1} & D_1 \ar[d]^{g_1}\\ G'_1 & D'_1 \ar[r]_{g'_1} \ar@{>->}[l]^{f'_1} & G_2}\]

	If a switch of $\dder{D}_0$ and $\dder{D}_1$ exists we say that they are \emph{switchable}. If $(\der{D}, \alpha, \omega)$ is a decorated derivation of length $2$, and $(i_0, i_1)$ an independence pair between $\dder{D}_0$ and $\dder{D}_1$, a \emph{decorated switch} of $(\der{D}, \alpha, \omega)$ along $(i_0, i_1)$ is $(\der{E}, \alpha, \omega)$, where $\der{E}$ is a switch for $\der{D}$ along $(i_0, i_1)$.
\end{definition}

\begin{remark}\label{rem:quadrati}
Let $\der{D}$ be a derivation made by two sequentially independence direct derivations with independence pair $(i_0, i_1)$. Suppose that $\der{E}$ is a pre-switch for $\der{D}$ along $(i_0, i_1)$. Then we have the solid part of the diagram below. If, moreover $\der{E}$ is a switch, then we can complete it with the dotted arrow $z_1\colon P\to D'_1$. 
		\[\xymatrix@C=15pt@R=15pt{&&R_0 \ar@/_.8cm/[ddrr]_(.2){i_0}|(.7)\hole
		\ar[dr]^{h_0}&& L_1\ar@/^.8cm/[ddll]^(.2){i_1}
		\ar[dl]_{m_1}\\&K_0\ar[dr]^{k_0}\ar@{>->}[dl]_{l_0}
		\ar@/_.6cm/[ddrr]|(.36)\hole^(.65){u_0}\ar[ur]^{r_0}&& G_1 &&
		K_1
		\ar@/^.6cm/[ddll]|(.36)\hole_(.65){u_1}\ar[dl]_{k_1}\ar@{>->}[lu]_{l_1}
		\ar[dr]^{r_1}\\L_0
		\ar@/_.6cm/[ddrr]_(.2){i'_1}|(.31)\hole|(.81)\hole
		\ar[dr]^(.4){m_0}|(.61)\hole && D_0
		\ar@{>->}[dl]|(.4)\hole_(.65){f_0}\ar[ur]|(.5)\hole^(.7){g_0}&&D_1
		\ar[dr]|(.4)\hole^(.65){g_1}
		\ar@{>->}[ul]|(.5)\hole_(.65){f_1}&&R_1\ar@/^.6cm/[ddll]^(.2){i'_0}|(.31)\hole|(.81)\hole\ar[dl]_{h_1}|(.6)\hole\\&G_0
		&&P\ar@{.>}[dr]_{z_1}	\ar@{>}[dl]^{z_0}\ar[ur]^(.4){p_1}\ar@{>->}[ul]_(.4){p_0}&&G_2\\L_1	\ar@/^.6cm/[uurr]^(.2){i_1} \ar[ur]_(.35){m'_0}|(.61)\hole && D'_0	\ar@{>->}[ul]|(.4)\hole_(.65){f'_0}\ar[dr]|(.5)\hole_(.65){g'_0}	&&D'_1\ar[ur]|(.4)\hole^(.65){g'_1} \ar@{>->}[dl]|(.5)\hole^(.65){f'_1}	&& R_0 \ar[ul]^(.35){h'_1}|(.61)\hole\ar@/_.6cm/[uull]_(.2){i_0}\\ &K_1	\ar[ur]_{k'_0} \ar[dr]_{r_1}	\ar@{>->}[ul]^{l_1}\ar@/^.6cm/[uurr]_(.65){u_1}&&G'_1&& K_0\ar[ur]_{r_0} \ar[ul]^{k'_1} \ar@{>->}[dl]^{l_0} \ar@/_.6cm/[uull]^(.65){u_0}\\&& R_1	\ar[ur]_{h'_0}\ar@/^.8cm/[uurr]^(.2){i'_0}&& L_0 \ar[ul]^{m'_1} \ar@/_.8cm/[uull]_(.2){i'_1} |(.69)\hole }\] 
\end{remark}

\begin{lemma}
Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation of length $2$ and $(i_0, i_1)$ an independence pair between $\dder{D}_0$ and $\dder{D}_1$. Given a pre-switch $\der{E}=\{\dder{E}_{i}\}_{i=0}^1$ for $\der{D}$ along $(i_0, i_1)$, the following are equivalent:
	\begin{enumerate}
		\item $\der{E}$ is a switch;
		\item if $(P, \{p_i\}_{i=0}^1)$ is a pullback for $f_{\der{D},1}$ along $g_{\der{D},0}$, then $\iota_{\der{E}, G_2} \circ g_1\circ p_1=\iota_{\der{E}, G'_1}\circ g'_0\circ z_0$.
	\end{enumerate}
\end{lemma}
\begin{proof}$(1 \Rightarrow 2)$. If $\der{E}$ is a switch then we have
	\begin{align*}
		\iota_{\der{E}, G_2} \circ g_1\circ p_1&=\iota_{\der{E}, G_2} \circ g'_1\circ z_1\\&=\iota_{\der{E}, D'_1}\circ z_1\\&=\iota_{\der{E}, G'_1}\circ f'_1\circ z_1\\&=\iota_{\der{E}, G'_1}\circ g'_0\circ z_0
	\end{align*}

	\smallskip \noindent $(2\Rightarrow 1)$. By hypothesis the solid part of the diagram below commutes.
	\[\xymatrix@C=30pt{P \ar[r]^{z_0} \ar@{.>}[dr]^{z_1}\ar[d]_{p_1} & D'_0 \ar[r]^{g'_0} & G'_1 \ar@{>->}[dr]^{\iota'_{1, G'_1}}\\ D_1 \ar[dr]_{g'_1}& D'_1 \ar[d]^{g'_1}\ar@{>->}[r]^{f'_1} &G'_1 \ar@{>->}[r]^{\iota'_{1, G'_1}} &\lpro \dder{E}_0 \rpro 	\ar@{>->}[d]^{p_1} \\ & G'_1 	\ar[rr]_{\iota_{\der{E}, G_2}}&& \tpro{E} }\]

	By \Cref{rem:zero1} the inner square is a pullback, proving that the dotted arrow $z_1\colon P\to D'_1$ exists.  Since, by \Cref{lem:colim} $\iota'_{1, G'_1}$ is in $\mathcal{M}$, we conclude that $f'\circ z_1=g'\circ z_0$ as wanted.
\end{proof}

We end this section showing that, if a switch exists, then any pre-switch can be turned into one composing it with an isomorphism.

\begin{proposition}\label{prop:pretosw} 	Let $\der{D}=\{\dder{D}_{i}\}_{i=0}^1$ be a derivation and suppose
	that both $\der{E}=\{\dder{E}_i\}_{i=0}^1$ and
	$\der{F}=\{\dder{F}_i\}_{i=0}^1$ are pre-switches of $\dder{D}_0$ and
	$\dder{D}_1$ along the same independence pair $(i_0, i_1)$. Let $\{\phi_X\}_{X\in \Deltamin(\der{E})}$ be the abstraction equivalence built in \Cref{thm:switch_uni}. If $\der{F}$ is a switch then also $\der{E}*\phi_{G_2}$ is one.
\end{proposition}
\begin{proof}
	contenuto...
\end{proof}

### 2025-03-21 full_version/permutations.tex
minori

\begin{proposition}\label{prop:unique} Let $\X$  be a $\mathcal{M}$-adhesive category. Suppose that the two direct derivations $\mathscr{D}$ and $\mathscr{D'}$ below, with the same match and applying the same left $\mathcal{M}$-linear rule $\rho$ are given.
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
	Then there exist isomorphisms $t\colon D\to D'$ and $s\colon H\to H'$ as in the following diagram.
	\[\xymatrix@C=40pt{&&D' \ar[r]^{g'} \ar@{>->}@/_.45cm/[dll]_{f'}&H'\\G & L \ar[l]_{m} & K \ar[u]_{k'} \ar[d]^{k}\ar[r]^{r} \ar@{>->}[l]_{l} &R\ar[u]^{h'} \ar[d]_{h}\\&&D\ar@{>->}@/^.45cm/[ull]^{f}\ar@/^.4cm/@{.>}[uu]^(.4){t}|\hole \ar[r]_{g}&H\ar@/_.4cm/@{.>}[uu]_{s}}\]
\end{proposition}
\begin{proof}
	By construction, the pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$ and $n$. Thus, the existence of the isomorphism $t\colon D\to D'$ follows from the second point of \Cref{lem:radj}. Now, computing we have
	\begin{align*}
		g'\circ t \circ k &= g' \circ k'\\&=h'\circ r
	\end{align*}
	Hence, we have the wanted $s\colon H\to H'$. To see that $s$ is an isomorphism, consider the diagram 
	\[\xymatrix{K  \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}& \ar[r]_{t} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{s}& H'}\]
	By hypothesis the whole rectangle and its left half are pushouts, therefore, by \Cref{lem:po1} its right square is a pushout too. The claim now follows from the fact that the pushout of an isomorphism is an isomorphism.
\end{proof}

>>>>

\begin{proposition}\label{prop:unique} Let $\X$  be a $\mathcal{M}$-adhesive category. Suppose that the two direct derivations $\mathscr{D}$ and $\mathscr{D'}$ below, with the same match and applying the same left $\mathcal{M}$-linear rule $\rho$ are given.
	\[\xymatrix{L \ar[d]_{m}& K \ar[d]^{k}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h} & L \ar[d]_{m}& K \ar[d]^{k'}\ar@{>->}[l]_{l} \ar[r]^{r} & R \ar[d]^{h'}\\G & \ar@{>->}[l]^{f} D \ar[r]_{g}& H & G & \ar@{>->}[l]^{f'} D' \ar[r]_{g'}& H'}\]
	Then there exist two unique isomorphisms $t\colon D\to D'$, $s\colon H\to H'$ as in the diagram below.
	\[\xymatrix@C=40pt{&&D' \ar[r]^{g'} \ar@{>->}@/_.45cm/[dll]_{f'}&H'\\G & L \ar[l]_{m} & K \ar[u]_{k'} \ar[d]^{k}\ar[r]^{r} \ar@{>->}[l]_{l} &R\ar[u]^{h'} \ar[d]_{h}\\&&D\ar@{>->}@/^.45cm/[ull]^{f}\ar@/^.4cm/@{.>}[uu]^(.4){t}|\hole \ar[r]_{g}&H\ar@/_.4cm/@{.>}[uu]_{s}}\]
\end{proposition}
\begin{proof}
	By construction, the pairs $(k, f)$ and $(k', f')$ are pushout complements of $l$ and $n$. Thus, existence and uniqueness of the isomorphism $t\colon D\to D'$ follows from the second point of \Cref{lem:radj}. Now, computing we have
	\begin{align*}
		g'\circ t \circ k &= g' \circ k'\\&=h'\circ r
	\end{align*}
	Hence, we have a unique arrow $s\colon H\to H'$. To see that $s$ is an isomorphism, consider the diagram 
	\[\xymatrix{K  \ar@/^.4cm/[rr]^{k'}\ar[d]_{r} \ar[r]_{k}& \ar[r]_{t} D \ar[d]^{g}& D' \ar[d]^{g'}\\ R \ar@/_.4cm/[rr]_{h'} \ar[r]^{h}& H \ar[r]^{s}& H'}\]
	By hypothesis the whole rectangle and its left half are pushouts, therefore, by \Cref{lem:po1} its right square is a pushout too. The claim now follows from the fact that the pushout of an isomorphism is an isomorphism.
\end{proof}
